Considering scale in multilayer animal social networks

### 2020-06-15 paper/manuscript.Rmd
temp

Our results highlight the influence of social, spatial and temporal scale on
multilayer networks. For social scale, we found individuals became more
connected with increasing spatial distance threshold. Increasing land cover
resolution lead to a decrease in availability of patchy and rare habitats and,
correspondingly, lower connectivity of individuals in these habitats. Temporal
windows highlighted clear periods of increased sociality across individuals.
Finally, increasing number of observations lead to a stabilizing in network
measures. As such, scale is inherently tied to the generation  of multilayer
networks and metrics evaluating them, as well as the individual layers that make
them up.

Monolayer network approaches have been increasingly used to link social network
position with individual fitness, including in primates [@Brent_2017;
@Thompson_2019], dolphins [@Stanton_2012; @Stanton_2011], and birds
[@McDonald_2007; @Royle_2012]. Social scale has emerged as an important variable
for understanding fitness consequences in the context of social network
positions <!-- TODO JWT check --> [@Almeling_2016; @Berger_2015; @Brent_2017;
@Holekamp_2011]. Defining the social scale at which to generate networks is an
essential consideration for social network analysis <!-- with direct impacts on
outcomes blah-->[@Castles_2014; @Carter_2015; @Farine_2015a]. Proximity-based
social networks generated using GPS location data require using a social scale,
specifically a spatial distance threshold, to define social associations. We
found increasing individual graph strength across three habitat layers with
increasing spatial distance threshold (Figure \@ref(fig:socres)). In addition,
habitat layers differed in their response to spatial distance threshold, with
open and forage have a much larger increase in graph strength and variation
across individuals. These results emphasize that multilayer network metrics are
directly influenced by social scale, in this case the spatial distance
threshold.

@@

Multilayer networks represents a novel framework for explicitly testing the
influence of habitat selection on sociality across scales. Animal social systems
can be parsed by spatial contexts, defined by habitat types as in
Figure \@ref(fig:lcres). Given that habitat data are available across a wide
range of spatial scales, we evaluated the influence of habitat spatial
resolution on multilayer network metrics. The results indicate varied influence
of increasing spatial resolution depending on the habitat layer. As spatial
resolution increased, edge overlap decreased in forage while remaining stable in
open and forest habitats (Figure \@ref(fig:socres) D). Forage is a rare and
patchy habitat on Fogo Island, as is indicated by a decrease in the proportion
of relocations in this habitat type with increasing spatial resolution (Figure
S1)<!--TODO-->. Similar to increasing spatial distance threshold, the influence
of increasing spatial resolution on graph strength was notably different between
habitat layers. Forest layers showed little change in graph strength with
increasing spatial resolution and variation between individuals. Forage and open
layers had greater individual variation in graph strength, while forage was the
sole layer to show large responses in graph strength to increasing spatial
resolution (Figure \@ref(fig:socres) E). As such, multilayer network measures
related to habitat are explicitly dependent on spatial scale. Edge overlap can
indicate if some associations between individuals are dependent on the
habitat-context [@Battiston_2014]. If the spatial resolution of the habitat data
are insufficient to capture rare habitats, context-specific relationships
between individuals may not be captured by multilayer networks.

The effect of seasonality of resources on sociality can be directly
measured by differences in association between individuals across time windows of resource availability. Using an ordinal multilayer network consisting of 20 discrete time windows throughout the study period, we observed periods of increased sociality
in caribou on Fogo Island (Figure \@ref(fig:temp))

<!-- These examples of explicit integrations of habitat -->
<!-- selection and sociality across social, spatial and temporal scales highlight the -->
<!-- potential for novel insights from multilayer networks. -->

@@

## Temporal scale
Temporal scale is also important for understanding the repercussions of social
phenotypes. Varying the time window length and position for caribou, we found
there was a lot of variation in the multilayer network metrics (Figure
\@ref(fig:winlen), Figure \@ref(fig:winpos)). These changes in the outcome
depending on how the time window is defined highlight how critical it is to
properly define the time window appropriately to address the questions and
hypotheses of interest. For instance, social phenotypes can change over ontogeny
in ways that can affect fitness [@Berger_2015; @Brent_2017]. Thus, it is
critical to carefully consider the temporal scale that is meaningful for the
proposed hypotheses. Further complicating the issue of selecting suitable time
scales across ontogeny, individuals progress through development at different
rates [@Tarka_2018]. This variation can make defining appropriate temporal
scales for testing the fitness implications of an individual’s social phenotype
over different stages of ontogeny even more difficult when they vary by
individual. However, considering fine temporal scales provide insight across
developmental stages that may otherwise be missed [@Turner_2017]. Defining
temporal scales is a challenge in multilayer networks just as in monolayer
networks; however, multilayer networks provide metrics that enable researchers
to look at similarities and differences in separate layers representing
different stages of ontogeny, or seasons as in our case study, which can be
difficult to detect.

Furthermore, multilayer networks provide a framework
to test hypotheses linking social phenotypes and fitness across temporal scales.
Because of the non-independent nature of social interactions and associations,
especially over time, it is hard to test hypotheses about social dynamics with a
suitable null control [@Farine_2017; @Proskurnikov_2017]. Using the framework of
multilayer networks, making layers of different temporal periods provides a way
to make appropriate null models to test questions about the dynamics of an
individual’s social network position over multiple aggregated time periods.

>>>>

Our results highlight the influence of social, spatial and temporal scale on
multilayer networks. For social scale, we found individuals became more
connected with increasing spatial distance threshold. Increasing land cover
resolution lead to a decrease in availability of patchy and rare habitats and,
correspondingly, lower connectivity of individuals in these habitats. Temporal
windows highlighted clear periods of increased sociality among individuals.
Finally, increasing number of observations lead to a general stabilizing in network
metrics. As such, scale is inherently tied to the generation  of multilayer
networks and metrics evaluating them, as well as the individual layers that make
them up.

Monolayer network approaches have been increasingly used to link social network
position with individual fitness, including in primates [@Brent_2017;
@Thompson_2019], dolphins [@Stanton_2012; @Stanton_2011], and birds
[@McDonald_2007; @Royle_2012]. Social scale has emerged as an important variable
for understanding fitness consequences in the context of social network
positions <!-- TODO JWT check --> [@Almeling_2016; @Berger_2015; @Brent_2017;
@Holekamp_2011]. Defining the social scale at which to generate networks is an
essential consideration for social network analysis <!-- with direct impacts on
outcomes blah-->[@Castles_2014; @Carter_2015; @Farine_2015a]. Proximity-based
social networks generated using GPS location data require using a social scale,
specifically a spatial distance threshold, to define social associations. We
found increasing individual graph strength across three habitat layers with
increasing spatial distance threshold (Figure \@ref(fig:socres)). In addition,
habitat layers differed in their response to spatial distance threshold, with
open and forage with a larger increase in graph strength and greater variation
across individuals. These results emphasize that multilayer network metrics are
directly influenced by social scale, in this case the spatial distance
threshold.

@@

Multilayer networks represent a novel framework for explicitly testing the
influence of habitat selection on sociality across scales. Animal social systems
can be parsed by spatial contexts, for example as defined by habitat types as in
Figure \@ref(fig:lcres). Given that habitat data are available across a wide
range of spatial scales, we evaluated the influence of habitat spatial
resolution on multilayer network metrics. The results indicate varied influence
of increasing spatial resolution depending on the habitat layer. As spatial
resolution increased, edge overlap decreased in forage while remaining stable in
open and forest habitats (Figure \@ref(fig:socres) D). Forage is a rare and
patchy habitat on Fogo Island, as is indicated by a decrease in the proportion
of relocations in this habitat type with increasing spatial resolution (Figure
S1)<!--TODO-->. Similar to increasing spatial distance threshold, the influence
of increasing spatial resolution on graph strength was notably different between
habitat layers. Forest layers showed little change in graph strength with
increasing spatial resolution and variation between individuals. Forage and open
layers had greater individual variation in graph strength, while forage was the
sole layer to show large responses in graph strength to increasing spatial
resolution (Figure \@ref(fig:socres) E). As such, multilayer network measures
related to habitat are explicitly dependent on spatial scale. Edge overlap can
indicate if some associations between individuals are dependent on the
habitat-context [@Battiston_2014]. If the spatial resolution of the habitat data
are insufficient to capture rare habitats, context-specific relationships
between individuals may not be captured by multilayer networks.

The effect of seasonality of resources on sociality can be directly measured by
differences in sociality in discrete time windows. Using an ordinal multilayer
network consisting of 20 discrete time windows throughout the study period, we
observed periods of increased sociality in caribou on Fogo Island (Figure
\@ref(fig:temp)). Edge overlap was higher between approximately October and May
(time windows 1, 6-11, 15-20) compared to between approximately May and October
(time windows 2-5, 12-15) (Figure \@ref(fig:temp) A).
<!-- not done -->

Furthermore, the selection of time windows is essential for both capturing life
stages of interest and having sufficient data to construct robust networks.
Social phenotypes can change over ontogeny in ways that can affect fitness
[@Berger_2015; @Brent_2017]. Further complicating the issue, individuals can
vary in their rate of development [@Tarka_2018]. This variation can make
defining the appropriate temporal scale for testing the fitness implications of
an individual’s social phenotype over different stages of ontogeny difficult.
However, considering fine temporal scales has been show to provide insight across developmental
stages that may otherwise be missed [@Turner_2017]. Furthermore, given the non-independent nature of social interactions and associations,
especially over time, it can be challenging to test hypotheses about social dynamics with a
suitable null control [@Farine_2017; @Proskurnikov_2017]. Using the framework of
multilayer networks, ordinal layers of discrete time windows provide a way
to make appropriate null models to test questions about the dynamics of an
individual’s social network position over multiple aggregated time periods. <!--hmm-->

These results
suggest that ordinal multilayer networks can effectively capture discrete time
windows of aggregation and dissaggregation related to resource availability, and
highlight the importance of carefully selecting the temporal scale of analysis.

@@

<!-- temp -->
These changes in the outcome
depending on how the time window is defined highlight how critical it is to
properly define the time window appropriately to address the questions and
hypotheses of interest.
